
Characters
-----------

Adam Olsen
+++++++++
 particle physics professor at University of Wisconsin-Madison
 husband of Kara Olsen
 talk at Cook County
  - they're not there to learn anything
   - *** later "you gave me a glimpse - picture metaphor" 
 works on upgrade set for 2013
  "or better yet", "the point is, was good like that, probably"

 
Kara Olsen 
++++++++++
 origin of life professor at UW Madison
 Wife of Adam Olsen
 runs, yoga, mornings
 undergrad Loyola

Alice Brown
++++++
 Adam's students
 outgoing
 Always liked Alice + Bob
 firstyear
 "that kind of thing","the ... type, wow, it turns out" "for sure" doubling words - "she wouldn't have to have to 
 informal style
 alice says the tricks physicists love are the ones that make it look less complicated than it really is
  - later: maybe our formulation just makes it look more complicated than it was (Nima)

Elsa Palmer
++++
 quiet
 best friend in early school Sylvia Butts
 - hawk playground

1st years

Daniel
++++++
 friend of Elsa 
 math, smart
 "at least" 

Nate
 - plasma
 - was in q1,2 statmech w/ alice, em first term
 - friendly collaborative
Harold
 - Harry
 - small, skinny, had been called harold before, now part of the gang, has a nickname
Ben
 - came up with nickname Harry
Joseph
Chris
Alex
 - wrote on board, neat handwriting



Steven Jacobs
++++++++++++
 Senior prof at UW Madison physics
 will be 76 in 2015

Twins
--------
Chloe + Walter (3 in 2008)
 - kids of Kara + Adam

cowboys


big speech
 - I am just a physicist and this is just a publicity stunt.
  - cowboy


- student who dropps out

- ugandan
  friend shows up late for test - one shot
  little sister had diarrhea
  - everyones little sister has dirrahea, you need to make a choice. education is our way out of here.

- tryharder girl

- cool girl
 - samantha?

- professors

- technitions

- russian

- with name like that he must be bengali

- Madison improv club
 vulgar word play

"The problem with the english language," the comedian said, "is that it's so confusing. Like if I want to negate something I can put un in front like undone or unknown but that doesn't always work. Take hard for example, the opposite isn't unhard, it's easy. And so people usually get the wrong impression when I walk into a room and say 'all of these dead bodies are making me uneasy.'"




%A particle collision at the LHC isn't just a particle collision. Or it isn't just one particle scattering nicely off of another.   



LHC Timeline
------------
1993 10 21 Superconducting Super Collider cancelled
1997 CMS and Atlas approved, US joins as observer
1998 Gallo-Roman ruins discovered at CMS site
2008 09 10 LHC first beam
2008 09 19 magnet incident
2009 04 30 last magnet back in 
2009 11 20 beams back 2.63 TeV
2010 03 30 7 TeV collisions
2011 10 18 almost 6 /fb
2012 04 05 8 TeV collisions
2012 07 04 higgs announcement
2013 02 16 LS1 begins
2013 10 08 Higgs/Englert Nobel prize
2015 06 03 13 TeV collisions
2016 02 11 LIGO gravitational waves


Story Timeline
--------------- 
2008  April starts
  - Alice + Elsa working for Adam Olsen on CSCs 
   - upgrade set for 2013
  first year students

2009 
 - do their second year
  - friends, etc
 Summer 2009, finish 2nd year, go to CERN in June

2010
 3 year
2011
 4 year
2012
 5 year - graduate on higgs
2013
 Postdoc 1
2014 
 Postdoc 2
2015


Talking with russian motorcyclist
-------------------------------
"Those are some serious pants"

He heard the strangers words, but didn't turn around, as if unsure if he was being addressed though he was sure he was being addressed. he was still wearing his heavy motorcycle pants and they were serious, but more importantly who could possibly be talking to him? He turned his head slightly and saw Alice looking him straight in the eyes. He turned the rest of the way in the chair and sat up just a bit straighter. Fuck. English.
"Thank you. Are my pants for motorcycle. Good on day like today." She didn't stop him, like she was paying attention. "I bought them in Ukraine. In Xyyalkcnklabl. Direct from manufacture. Less than forty percent from sell price. Very tough. Sew four layer consecutive." He stopped. She had understood right? He knew his accent could be too thick for those westerners sometimes. Why did he have such a stupid grin on his face?

"Consecutive. Nice word." 

Fuck, what? Women weren't like that in Russia. In Mother Russia. When he was fifteen he and his friends would toast To Mother Russia in english, exaggerating their already thick accents and enjoying the irony like any good russian. "It is a simple word. Everybody knows this word. Consecutive. Is like counting."

random ideas
--------------
 engineer likes thinkpad, that little red knob for mouse control and hacks it so that's his bluetooth mouse
 vertical screen
 


oneliners
-----------
there are 24 hours in the day, and then there's the night
bright future
USB drives are spin-1/2 particles 
 ah
 do you get it? you keep rotating them, 180, 360 doesn't work, because it's somehow flipped over and you have to go through 720 to get it to fit back in?
 ah, i see. what's a US B-drive?
 like for a computer. a memory stick.
 ohhh, a USB drive. ah yes. they are difficult to orient. i see. hahah. that is a good joke. i was thinking about cars and how in the US they use automatic transmission, maybe they have a B-drive. 


scientist optimist for future
-----------------------
to be a scientist is to be an optimist. or at least to be hopeful about the future. the only context in which science, pure science, or math for that matter, or come to think of it, there's a whole range of fields of study for which this is true but pure science and math especially, the only context in which they make any sense is one where you assume that there is a future and that the people of the future are free to learn and study and ask the most difficult questions about what is really going on. 

He continued, "Building on previous knowledge is the only way science can grow. it's the only way society can grow. really, a measure of the science a society understands is an assessment of that society's progress." 
 - you don't think science can flourish under repression?
no because what's the point? there is no point. the only point is to learn. to understand more. repression needs violence and violence needs justification. since the real reason why anyone does science has no justification, it can't possibly be a central value to a repressive government. and things that aren't central to repressive governments get cut. especially, he added, when they're in the business of asking questions. 
 ++ chatted about something else
 - you don't think science has justifications beyond this pure ideal of societies growing and people interrogating the universe? 



Plaque on pump room door - passover dinner?
----------------------
 israel palestine peace 
 there were jews in cms - not a ton, and mostly american 
  - didn't really stand out, most were atheist jews 
  - many practicing jews too, passover dinner party
   - not in your face about it like christians
 you didn't really notice that many jews in europe. oh yea, fuck. that was fucked up
  - i wonder if the fact that jews weren't in your face about it is what made the christians feel so threatened



doc didn't have phone, needs to give report at RC meeting about it
----------------------------
"Lesson #1 - have your phone handy"
"I don't understand, these phones should all be handy"
"ah yes, I just mean they should have it with them, not that it's a landline"
"yes but the point of a handy phone is that you keep it handy, I keep mine in my pocket"
"I agree, this isn't a new idea, it's just good advice"
"you mean advice for yourself"
"yes"


french/american comparison - cart before horse
-------------------
 frenchman - you americans put the cart before the horse
 american - no, see that's the problem with the french, you need to wait for the horse before you even start thinking about the cart. For an american, we say, we'll need both the cart and the horse to get the job done, so let's work on getting both right now. And since you're already building me a cart, how about you tell me how much more it would cost if I say I wanted two.


She checked her email
------------------
Re: Draft in response to the referee comments
Re: spurious streams in the transfer manager
Re: Impact plots - error
Daily Run Meeting Summary
crab fails with error 139 but cmsRun works locally
Re: New electron ID available in standard CMSSW modules
Re: [Indico][Event Reminder] MC \& interpretation (Even Week) (date, 16:00 Europe/Zurich)
Re: Negative Signal
Re: crab fails with error 139 but cmsRun works locally
\\xi symbol not printed correctly
Re: crab fails with error 139 but cmsRun works locally
response letter v1
Re: crab fails with error 139 but cmsRun works locally
Fwd: Arxiv label on additional figures in webpages
MET meeting - DATE
Fwd: LHC Plans for the night
Goodness of fit for multi signal models
Re: Goodness of fit for multi signal models
Re: Goodness of fit for multi signal models
Re: Goodness of fit for multi signal models
Re: crab fails with error 139 but cmsRun works locally
Re: Goodness of fit for multi signal models
Re: Goodness of fit for multi signal models
Development Release of CMSSW_XXXXXXX Now Available
Re: accessing jet energy scale uncertainties





Scary Crows
----------
Going to CMS week in NYC (NYU, Columbia?)
 15 people at lunch table, maybe knew half
 asked about NYC, gets phone number for "mayor of statin island", call on thursday night to plan for friday
 - 518 area code, where was that
 - decision about if she should call, while there meets mayor (secretary)
  - would you like to go for an adventure?
   - friend picks her up on bike trailer (dennis)
    ( falls apart? meet up with red range rover)
   - had been told she might come and call
   
  - lives in mansion, music studio, bones in backyard from gangsters
  - zach dennis logan mike
  - party, costumes, meet other girls 



IceCube
--------
Friend works for ice cube
 - wants to go to south pole
  - giant orgy
  - would even be a space janitor


Adam Talkin with indian at cern
-------------------------
i have a friend from near bollywood
and where is that
bombay
he smiled - if your friend is from bombay then he may be from bollywood itself. bollywood comes from bombay hollywood and it is not a place, but neither is bombay. it is now called mumbai. so it could be said that bollywood and bombay are now even more the same than ever, neither existing but both claiming quite a few residents. do you know the problem with american movies, he added, not enough singing.

political correctness
-------------------
kids at school called him retard boy 
 - yea kids suck, they're mean
at least they say what they're thinking, they stopped calling me retard boy in high school but they never stopped thinking it


convo
---------
 has frayed cord for phone on bed
 dude, you're going to die in a fire
 i would wake up
 true, you'd have to be pretty drunk, but you'd still burn down your matress, or put a hole in it
 it's fine
 it's still a bad idea, can i put electrical tape on it




Alice Intro appendix
----------------
solve problems
q4: 
Bob, yes, in a way, no, no, no, no, escaping mortality, yes, no, not literally at least.
Quantum2, ps5

french translated
----------------
Hey it goes?
 - It goes. It goes?
It goes.



alice intro ref
---------------
even difficult problem of mapping concepts onto complicated reality wasn't really reality
it's mapping it onto a still idealized universe that only consists of this particle and this potential
no people, no milky way, not even an infinite frictionless pond

back of the book
-------------
About the book:
give histogram showing word frequency for most common words
metadata
